---
title: 'Figure: Species Ranking by Engraftment'
author: "Sebastian Schmidt"
date: "2021-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(stringsAsFactors = FALSE)
```

## Prepare Environment

```{r Load Libraries, echo=F, warning=FALSE}
library("tidyverse", warn.conflicts=F, quietly=T)
```

Set file names and basic parameters.

```{r Set Parameters}
#Preallocate global data structures
PARAM <- list()

#Set relevant folder names
PARAM$folder.Rmd <- paste0(getwd(), "/")
PARAM$folder.base <- gsub("Rmd/", "", PARAM$folder.Rmd)

load(paste0(PARAM$folder.base, "data/param.analysis_transmission.Rdata"))
```

Load data.

```{r}
#load(PARAM$file.sample)
load(PARAM$file.sample)
load(PARAM$file.phenotype)
load(PARAM$file.data_transmission)
load(PARAM$file.data_transmission.scored)
load(PARAM$file.data_transmission.scored.consolidated)
load(PARAM$file.mOTU)
load(PARAM$file.mOTU.ANI)
load(PARAM$file.tree)
load(PARAM$file.beta_div)
load(PARAM$file.allele_distances)
load(PARAM$file.allele_diversity)
load(PARAM$file.gmm_gmgc)
load(PARAM$file.gmm_species)
load(PARAM$file.sample)

#LASSO results
load(paste0(PARAM$folder.results, "results.engraftment.LASSO.Rdata"))
#OLS regression results
load(paste0(PARAM$folder.results, "results.engraftment.OLS.Rdata"))
```

Reset some parameters

```{r}
PLOT$point.size <- 0.5

#Re-set outcome colours
PLOT$colour$outcome$species_lost <- "#ffffb3"
PLOT$colour$outcome$species_retained <- "#FFE206"
PLOT$colour$outcome$rejection_consp <- "#FFFF33"
PLOT$colour$outcome$coexist.rec <- "#FDBF6F"
PLOT$colour$outcome$coexist.donor <- "#ff7f00"
PLOT$colour$outcome$engraft_consp <- "#A6CEE3"
PLOT$colour$outcome$engraft_novel <- "#1F78B4"
PLOT$colour$outcome$influx_consp <- "#DE77AE"
PLOT$colour$outcome$influx_novel <- "#C51B7D"
PLOT$colour$outcome$rejection_novel <- "#DEEBF7"
```

## Plot AUCs on Outcome Predictability

```{r, fig.width=10, fig.height=4}
collect.res.engraft.lasso.metric %>%
  filter(resp.var %in% c("oc.donor_takeover", "oc.rec_resilience", "oc.rec_turnover")) %>%
  filter(non.trivial>0.2) %>%
  filter(metric == "auc") %>% #group_by(resp.var, var.type) %>% tally() %>% arrange(desc(n))
  select(-metric, -non.trivial, auc=avg) %>%
  mutate(auc = ifelse(auc<0.5, 0.5, auc)) %>%
  separate(var.type, into = c("scope", "type"), sep = "\\.", extra = "merge") %>%
  #filter(scope == "ex_ante") %>%
  mutate(
    type = fct_relevel(type, c(
      "combined",
      "technical",
      "community.tax", "community.metabolic",
      "abd_pre", "abd_pre.facilitation", "abd_pre.exclusion",
      "abd_donor", "abd_donor.facilitation", "abd_donor.exclusion",
      "abd_post", "abd_post.pos", "abd_post.neg",
      "focal_species",
      "engraftment"
    )),
    var.group = case_when(
      type == "combined" ~ "full",
      type %in% c("community.tax", "community.metabolic") ~ "community",
      grepl("abd_pre", type) ~ "abd_pre",
      grepl("abd_donor", type) ~ "abd_donor",
      grepl("abd_post", type) ~ "abd_post",
      type %in% c("technical", "focal_species", "engraftment") ~ as.character(type)
    ),
    var.group = fct_relevel(var.group, c("full", "technical", "community", "abd_pre", "abd_donor", "abd_post", "focal_species", "engraftment")),
    resp.var = fct_relevel(resp.var, c("oc.rec_resilience", "oc.donor_takeover", "oc.rec_turnover"))
  ) %>%
  ggplot(aes(x = type, y = auc, color = resp.var, fill = resp.var, group = interaction(resp.var, type), alpha = auc)) +
    geom_hline(yintercept = c(0.5, 0.6, 0.7, 0.8, 0.9, 1), colour = "#969696", lwd = 0.2) +
    #Add jittered points
    geom_point(position = position_jitterdodge(jitter.width = 0.1), size = 0.4) +
    #Add boxplot
    geom_boxplot(outlier.colour = NA, lwd = 0.2, alpha = 1) +
    stat_summary(fun = median, geom = "crossbar", width = 0.75, colour = "#ffffff", lwd = 0.4, position = position_dodge()) +
    stat_summary(fun = median, geom = "point", colour = "#ffffff", size = 1, shape = 1, position = position_dodge(width = 0.75)) +
    scale_alpha_continuous(range = c(0.1, 0.9), limits = c(0.5, 1)) +
    scale_colour_manual(values = c(
      PLOT$colour$outcome$species_retained,
      #PLOT$colour$outcome$engraft_consp,
      PLOT$colour$outcome$engraft_novel,
      PLOT$colour$outcome$influx_novel
    )) +
    scale_fill_manual(values = c(
      PLOT$colour$outcome$species_retained,
      #PLOT$colour$outcome$engraft_consp,
      PLOT$colour$outcome$engraft_novel,
      PLOT$colour$outcome$influx_novel
    )) +
    facet_grid(. ~ scope + var.group, scales = "free_x", space = "free_x") +
    theme_minimal() +
    theme(
      axis.title = element_blank(),
      axis.text = element_blank(),
      panel.grid = element_blank(),
      strip.text = element_blank(),
      plot.margin = margin(r = 0, l = 0, t = 0, b = 0),
      panel.spacing.x = unit(0.05, "cm"),
      legend.position = "none"
    ) -> p.auc_boxplot
p.auc_boxplot

ggsave(p.auc_boxplot, filename = paste0(PARAM$folder.results, "engraftment.lasso_species.auc_boxplots.pdf"), width=13, height=2.5, useDingbats = F, units = "cm")
```


## Plot as network(s)

First, define set of species and set of variables to include.

```{r}
collect.res.engraft.lasso.metric %>%
  filter(resp.var %in% c("oc.rec_turnover", "oc.donor_takeover") & var.type == "ex_ante.combined") %>%
  filter(non.trivial>0.2) %>%
  group_by(species) %>% filter(n() > 1) %>%
  pull(species) %>% unique -> use.species.lasso

collect.res.engraft.lasso.coef %>%
  filter(species %in% use.species.lasso) %>%
  filter(var.type == "ex_ante.combined") %>%
  filter(resp.var %in% c("oc.rec_turnover", "oc.donor_takeover")) %>%
  filter(coef.nonzero > 0.1) %>%
  filter(!grepl("abd_pre", term) & !grepl("abd_donor", term)) %>%
  #filter(! term %in% unlist(v)) %>%
  group_by(term) %>% tally() %>% arrange(desc(n))
```

```{r}
v <- list()

v[["technical"]] <- c(
  #"min_depth",
  "indication.rcdi",
  "indication.mets",
  "antibiotics.summary",
  "fmt_route.nasal"
  #"fmt_sample_prep.fresh"
  #"bowel.prep.summary"
)

v[["community.tax"]] <- c(
  "dissimilarity.donor_pre",
  "richness.pre.scaled",
  #"richness.donor.scaled"
  "fr.pre.scaled"
  #"fr.donor.scaled"
)

v[["community.metabolic"]] <- c(
  "dist_gmm.donor_pre",
  "butyrate_production.pre",
  "butyrate_production.donor"
  #"proteolysis.pre",
  #"proteolysis.donor"
  #"saccharolysis.pre",
  #"saccharolysis.donor"
)

collect.res.engraft.lasso.coef %>%
  filter(species %in% use.species.lasso) %>%
  filter(var.type == "ex_ante.combined") %>%
  filter(resp.var %in% c("oc.rec_turnover", "oc.donor_takeover")) -> tmp.res

tmp.res %>%
  filter(grepl("abd_pre", term)) %>%
  #filter(coef.nonzero > 0.1) %>%
  group_by(term) %>%
  summarise(vip = sum(abs(coef.nonzero))) %>%
  arrange(desc(vip)) %>%
  mutate(mOTU = str_remove(term, "\\.abd_pre")) %>%
  left_join(data.taxonomy.gtdb %>% select(mOTU, family, label), by = "mOTU") %>%
  pull(term) %>% head(30) -> v[["abd.pre"]]

tmp.res %>%
  filter(grepl("abd_donor", term)) %>%
  #filter(coef.nonzero > 0.1) %>%
  group_by(term) %>%
  summarise(vip = sum(abs(coef.nonzero))) %>%
  arrange(desc(vip)) %>%
  mutate(mOTU = str_remove(term, "\\.abd_donor")) %>%
  left_join(data.taxonomy.gtdb %>% select(mOTU, family, label), by = "mOTU") %>%
  pull(term) %>% head(10) -> v[["abd.donor"]]

data.taxonomy.gtdb %>%
  filter(mOTU %in% c(str_remove(v[["abd.pre"]], "\\.abd_pre"), str_remove(v[["abd.donor"]], "\\.abd_donor"))) %>% #group_by(order) %>% tally
  mutate(
    order = fct_relevel(order, c("Bacteroidales", "Coriobacteriales", "Enterobacterales", "Lactobacillales", "Oscillospirales", "Lachnospirales", "Peptostreptococcales")),
    mOTU = fct_relevel(mOTU, c("ref_mOTU_v2_0898", "ref_mOTU_v2_0899"))
  ) %>%
  arrange(order, family, mOTU) %>% pull(label) %>% as.character() -> order.species.lasso_abd

v[["focal.species"]] <- c(
  "abd_ratio.donor_pre",
  "allele_dist.donor_pre",
  "allele_div.pre",
  "allele_div.donor",
  "phylo.complementarity"
  #"ko.complementarity"
)
```

Pre-set graph layouts.

```{r}
x.max <- 200
x.increment <- 1

node.size_min <- 1.5
node.size_max <- 6
```

```{r}
data.taxonomy.gtdb %>%
  filter(mOTU %in% use.species.lasso) %>%
  mutate(
    taxonomy.summarised = case_when(
      genus %in% c("Bacteroides", "Phocaeicola") ~ "Bacteroides sp.",
      order == "Bacteroidales" ~ "other Bacteroidales",
      phylum %in% c("Proteobacteria") ~ "other",
      class %in% c("Bacilli", "Negativicutes") ~ "Bacilli_Negativicutes",
      order %in% c("Actinomycetales", "Coriobacteriales") ~ "Actinobacteria",
      grepl("Blauti", genus) ~ "Blautia sp.",
      genus %in% c("Faecalibacterium") ~ "Faecalibacterium sp.",
      family == "Oscillospiraceae" ~ "Oscillospiraceae sp.",
      family %in% c("Lachnospiraceae", "Ruminococcaceae") ~ paste("other", as.character(family)),
      class %in% c("Clostridia", "Clostridia_A") ~ "other Clostridia",
      #order %in% c("Peptostreptococcales", "Monoglobales_A", "Christensenellales", "Oscillospirales", "Lachnospirales") ~ "other Clostridia",
      TRUE ~ "other"
    ),
    taxonomy.summarised = fct_relevel(taxonomy.summarised, c(
      "Bacteroides sp.",
      "other Bacteroidales",
      "Actinobacteria",
      "Bacilli_Negativicutes",
      "Oscillospiraceae sp.",
      "Blautia sp.",
      "other Lachnospiraceae",
      "Faecalibacterium sp.",
      "other Ruminococcaceae",
      "other Clostridia",
      "other"
    )),
    genus = fct_relevel(genus, "Phocaeicola")
  ) %>%
  arrange(taxonomy.summarised, family, genus) %>%
  select(mOTU, label, taxonomy.summarised) %>%
  #Add indices for layout
  mutate(
    x.pos = row_number() * ((x.max - 11) / length(use.species.lasso)) + ((as.numeric(taxonomy.summarised) - 1) * x.increment)
  ) -> tmp.taxonomy

tibble(term = unlist(v)) %>%
  mutate(
    type = case_when(
      term %in% v[["technical"]] ~ "technical",
      term %in% v[["community.tax"]] ~ "community.tax",
      term %in% v[["community.metabolic"]] ~ "community.metabolic",
      term %in% v[["abd.pre"]] ~ "abd.pre",
      term %in% v[["abd.donor"]] ~ "abd.donor",
      term %in% v[["focal.species"]] ~ "focal.species"
    ),
    type = fct_relevel(type, names(v)),
    scope = case_when(
      term %in% c(v[["abd.pre"]], "antibiotics.summary", "bowel.prep.summary", "richness.pre.scaled", "fr.pre.scaled", "butyrate_production.pre", "allele_div.pre") ~ "recipient",
      grepl("indication", term) ~ "recipient",
      term %in% c(v[["abd.donor"]], "fmt_route.nasal", "richness.donor.scaled", "butyrate_production.donor", "allele_div.donor") ~ "donor",
      TRUE ~ "donor_rec"
    )
  ) %>%
  mutate(
    term_name = str_remove(string = str_remove(term, "\\.abd_donor"), "\\.abd_pre")
  ) %>%
  left_join(data.taxonomy.gtdb %>% select(mOTU, label), by = c("term_name" = "mOTU")) %>%
  mutate(term_name = ifelse(!is.na(label), label, term_name)) %>%
  select(-label) %>%
  mutate(
    term = fct_relevel(term, c(v[["technical"]], v[["community.tax"]], v[["community.metabolic"]], v[["abd.pre"]], v[["abd.donor"]], v[["focal.species"]])),
    term_name = fct_relevel(term_name, c(v[["technical"]], v[["community.tax"]], v[["community.metabolic"]], order.species.lasso_abd, v[["focal.species"]]))
  ) %>%
  arrange(type, term_name) %>%
  mutate(
    x.pos = row_number() * ((x.max - (4 * 6)) / 55) + ((as.numeric(type) - 1) * x.increment * 4)
  ) -> tmp.terms
tmp.terms
```

### Donor takeover data

```{r}
require(tidygraph)
require(ggraph)

tmp.res %>%
  filter(term %in% unlist(v)) %>%
  filter(resp.var == "oc.donor_takeover") %>%
  filter(coef.nonzero >= 0.1) %>%
  select(-var.type) %>%
  dplyr::mutate(
    direction = ifelse(coef.mean < 0, "neg", "pos"),
    coef = abs(coef.mean)
  ) %>%
  dplyr::transmute(
    from = term,
    to = species,
    weight1 = coef,
    weight2 = coef.nonzero,
    edge.type = direction
  ) %>%
  arrange(from, to) %>%
  as_tbl_graph(directed = T) %>%
  activate(nodes) %>%
  left_join(tmp.taxonomy %>% select(mOTU, label), by = c("name" = "mOTU")) %>%
  left_join(tmp.terms %>% select(term, term_name, type, scope), by = c("name" = "term")) %>%
  mutate(
    category = case_when(
      !is.na(type) ~ as.character(type),
      !is.na(label) ~ "species"
    ),
    label = ifelse(is.na(label), as.character(term_name), label)
  ) %>%
  select(-term_name) %>%
  mutate(
    var.type = type,
    type = name %in% use.species.lasso,
    degree_out = centrality_degree(),
    degree_in = centrality_degree(mode = "in"),
    degree = pmax(degree_in, degree_out)
  ) %>%
  activate(edges) %>%
  mutate(
    var.type = .N()$var.type[from]
  ) %>%
  filter(weight1 > 0) -> tmp.graph.colonisation
tmp.graph.colonisation %>% activate(nodes) %>% as_tibble() %>% arrange(var.type)
```

Manually layout the graph: colonisation network.

```{r, fig.height=6, fig.width=8}
#Create a fake layout first
(layout <- create_layout(tmp.graph.colonisation, "bipartite"))

#Manually re-layout
layout %>%
  left_join(tmp.taxonomy %>% select(mOTU, x.pos), by = c("name" = "mOTU")) %>%
  mutate(x = ifelse(!is.na(x.pos), x.pos, x)) %>% select(-x.pos) %>%
  left_join(tmp.terms %>% select(term, x.pos), by = c("name" = "term")) %>%
  mutate(x = ifelse(!is.na(x.pos), x.pos, x)) %>% select(-x.pos) -> layout

layout %>% arrange(x) %>% filter(category == "species")

#Plot graph
ggraph(tmp.graph.colonisation, x = layout$x, y = layout$y) +
  geom_edge_diagonal(strength = 1, aes(colour = edge.type, alpha = ..index.., edge_width = weight1)) +
  scale_edge_alpha('Edge direction', guide = 'edge_direction', range = c(0.05, 0.6)) +
  scale_edge_width_continuous(range = c(0.5, 2.5)) +
  #scale_edge_colour_distiller(palette = "PRGn", direction = 1) +
  scale_edge_colour_manual(values = c("#c51b7d", "#7fbc41")) +
  #geom_node_point(aes(size = degree, colour=scope), alpha = 1) +
  geom_node_point(aes(fill=scope, size=degree), colour="#bdbdbd", shape=21, stroke = 0.2, alpha = 1) +
  scale_size_continuous(range = c(node.size_min, node.size_max)) +
  #scale_colour_brewer(palette = "Paired", direction = 1) +
  scale_fill_manual(values = c(
    PLOT$colour$outcome$engraft_consp,
    "#238B45",
    PLOT$colour$outcome$species_retained
    ),
    #na.value = PLOT$colour$outcome$engraft_novel
    na.value = "#969696"
  ) +
  #facet_edges(~ var.type, ncol = 3) +
  theme_graph(
    #base_family = PLOT$font$base_family,
    #base_size = PLOT$font$size.base
  ) +
  theme(
    strip.text = element_blank(),
    plot.margin = margin(r = 0, l = 0, t = 0, b = 0),
    panel.spacing.x = unit(0.05, "cm"),
    panel.spacing.y = unit(0.05, "cm"),
    panel.background = element_blank(),
    legend.position = "none"
  ) -> p.graph_colonisation
p.graph_colonisation

p.graph_colonisation +
  scale_size_continuous(range = c(0.5 * node.size_min, 0.5 * node.size_max)) +
  scale_edge_width_continuous(range = c(0.25, 2)) +
  facet_edges(~ var.type, ncol = 2) -> p.graph_colonisation.faceted
p.graph_colonisation.faceted

ggsave(p.graph_colonisation, filename = paste0(PARAM$folder.results, "engraftment.lasso_species.graph_colonisation.pdf"), width=18, height=3, useDingbats = F, units = "cm")
ggsave(p.graph_colonisation.faceted, filename = paste0(PARAM$folder.results, "engraftment.lasso_species.graph_colonisation.faceted.pdf"), width=18, height=4.5, useDingbats = F, units = "cm")
```

### Turnover network

Prepare graph data.

```{r}
require(tidygraph)
require(ggraph)

tmp.res %>%
  filter(term %in% unlist(v)) %>%
  filter(resp.var == "oc.rec_turnover") %>%
  filter(coef.nonzero >= 0.1) %>%
  select(-var.type) %>%
  dplyr::mutate(
    direction = ifelse(coef.mean < 0, "neg", "pos"),
    coef = abs(coef.mean)
  ) %>%
  dplyr::transmute(
    from = term,
    to = species,
    weight1 = coef,
    weight2 = coef.nonzero,
    edge.type = direction
  ) %>%
  arrange(from, to) %>%
  as_tbl_graph(directed = T) %>%
  activate(nodes) %>%
  left_join(tmp.taxonomy %>% select(mOTU, label), by = c("name" = "mOTU")) %>%
  left_join(tmp.terms %>% select(term, term_name, type, scope), by = c("name" = "term")) %>%
  mutate(
    category = case_when(
      !is.na(type) ~ as.character(type),
      !is.na(label) ~ "species"
    ),
    label = ifelse(is.na(label), as.character(term_name), label)
  ) %>%
  select(-term_name) %>%
  mutate(
    var.type = type,
    type = name %in% use.species.lasso,
    degree_out = centrality_degree(),
    degree_in = centrality_degree(mode = "in"),
    degree = pmax(degree_in, degree_out)
  ) %>%
  activate(edges) %>%
  mutate(
    var.type = .N()$var.type[from]
  ) %>%
  filter(weight1 > 0) -> tmp.graph.turnover
tmp.graph.turnover
```

Manually layout the graph: colonisation network.

```{r, fig.height=6, fig.width=8}
#Create a fake layout first
(layout <- create_layout(tmp.graph.turnover, "bipartite"))

#Manually re-layout
layout %>%
  left_join(tmp.taxonomy %>% select(mOTU, x.pos), by = c("name" = "mOTU")) %>%
  mutate(x = ifelse(!is.na(x.pos), x.pos, x)) %>% select(-x.pos) %>%
  left_join(tmp.terms %>% select(term, x.pos), by = c("name" = "term")) %>%
  mutate(x = ifelse(!is.na(x.pos), x.pos, x)) %>% select(-x.pos) -> layout

#Plot graph
ggraph(tmp.graph.turnover, x = layout$x, y = layout$y) +
  geom_edge_diagonal(strength = 1, aes(colour = edge.type, alpha = ..index.., edge_width = weight1)) +
  scale_edge_alpha('Edge direction', guide = 'edge_direction', range = c(0.05, 0.6)) +
  scale_edge_width_continuous(range = c(0.5, 2.5)) +
  #scale_edge_colour_distiller(palette = "PRGn", direction = 1) +
  scale_edge_colour_manual(values = c("#c51b7d", "#7fbc41")) +
  #geom_node_point(aes(size = degree, colour=scope), alpha = 1) +
  geom_node_point(aes(fill=scope, size=degree), colour="#bdbdbd", shape=21, stroke = 0.2, alpha = 1) +
  scale_size_continuous(range = c(node.size_min, node.size_max)) +
  #scale_colour_brewer(palette = "Paired", direction = 1) +
  scale_fill_manual(values = c(
    PLOT$colour$outcome$engraft_consp,
    "#238B45",
    PLOT$colour$outcome$species_retained
    ),
    na.value = PLOT$colour$outcome$influx_novel
  ) +
  #facet_edges(~ var.type, ncol = 3) +
  theme_graph(
    #base_family = PLOT$font$base_family,
    #base_size = PLOT$font$size.base
  ) +
  theme(
    strip.text = element_blank(),
    plot.margin = margin(r = 0, l = 0, t = 0, b = 0),
    panel.spacing.x = unit(0.05, "cm"),
    panel.spacing.y = unit(0.05, "cm"),
    panel.background = element_blank(),
    legend.position = "none"
  ) -> p.graph_turnover
p.graph_turnover

p.graph_turnover +
  scale_size_continuous(range = c(0.5 * node.size_min, 0.5 * node.size_max)) +
  scale_edge_width_continuous(range = c(0.25, 2)) +
  facet_edges(~ var.type, ncol = 2) -> p.graph_turnover.faceted
p.graph_turnover.faceted

ggsave(p.graph_turnover, filename = paste0(PARAM$folder.results, "engraftment.lasso_species.graph_turnover.pdf"), width=18, height=3, useDingbats = F, units = "cm")
ggsave(p.graph_turnover.faceted, filename = paste0(PARAM$folder.results, "engraftment.lasso_species.graph_turnover.faceted.pdf"), width=18, height=4.5, useDingbats = F, units = "cm")
```



----


