---
title: 'Figure: Species Ranking by Engraftment'
author: "Sebastian Schmidt"
date: "2021-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(stringsAsFactors = FALSE)
```

## Prepare Environment

```{r Load Libraries, echo=F, warning=FALSE}
library("tidyverse", warn.conflicts=F, quietly=T)
```

Set file names and basic parameters.

```{r Set Parameters}
#Preallocate global data structures
PARAM <- list()

#Set relevant folder names
PARAM$folder.Rmd <- paste0(getwd(), "/")
PARAM$folder.base <- gsub("Rmd/", "", PARAM$folder.Rmd)

load(paste0(PARAM$folder.base, "data/param.analysis_transmission.Rdata"))
```

Load data.

```{r}
#load(PARAM$file.sample)
load(PARAM$file.sample)
load(PARAM$file.phenotype)
load(PARAM$file.data_transmission)
load(PARAM$file.data_transmission.scored)
load(PARAM$file.data_transmission.scored.consolidated)
load(PARAM$file.mOTU)
load(PARAM$file.mOTU.ANI)
load(PARAM$file.tree)
load(PARAM$file.beta_div)
load(PARAM$file.allele_distances)
load(PARAM$file.allele_diversity)
load(PARAM$file.gmm_gmgc)
load(PARAM$file.gmm_species)
load(PARAM$file.sample)

#LASSO results
load(paste0(PARAM$folder.results, "results.engraftment.LASSO.Rdata"))
#OLS regression results
load(paste0(PARAM$folder.results, "results.engraftment.OLS.Rdata"))
```

Reset some parameters

```{r}
PLOT$point.size <- 0.5

#Re-set outcome colours
PLOT$colour$outcome$species_lost <- "#ffffb3"
PLOT$colour$outcome$species_retained <- "#FFE206"
PLOT$colour$outcome$rejection_consp <- "#FFFF33"
PLOT$colour$outcome$coexist.rec <- "#FDBF6F"
PLOT$colour$outcome$coexist.donor <- "#ff7f00"
PLOT$colour$outcome$engraft_consp <- "#A6CEE3"
PLOT$colour$outcome$engraft_novel <- "#1F78B4"
PLOT$colour$outcome$influx_consp <- "#DE77AE"
PLOT$colour$outcome$influx_novel <- "#C51B7D"
PLOT$colour$outcome$rejection_novel <- "#DEEBF7"
```

## Species Ranking

Plot engraftment per species, ranked by "successful" engraftment across scored FMTs.

### Prepare data

Pre-compute summarised engraftment data.

```{r}
#Subset engraftment results to species scored in >= 50 FMTs
dat.transmission.scored.outcome %>%
  group_by(fmt.id) %>% top_n(1, timepoint.fmt) %>% ungroup() %>%
  filter(species %in% use.species.engraftment.gt_50) -> dat.transmission.scored.outcome.subset

#Summarise FMT outcomes
dat.transmission.scored.outcome.subset %>%
  dplyr::select(species, frac.don:influx_novel, inc.post) %>%
  pivot_longer(frac.don:influx_novel, names_to = "type") %>%
  group_by(species, type) %>%
  dplyr::summarise(
    n.FMT = n(),
    val = sum(value) / sum(inc.post)
  ) %>%
  ungroup() %>%
  pivot_wider(names_from = "type", values_from = "val") %>%
  arrange(frac.rec) -> dat.transmission.scored.outcome.subset.summarised

#Summarise frequency of binarised outcome scenarios per species
dat.transmission.scored.outcome.subset %>%
  dplyr::select(species, oc.donor_takeover:oc.rec_turnover) %>%
  group_by(species) %>%
  dplyr::mutate(n.FMT = 1) %>%
  summarise_all(sum) %>%
  ungroup() %>%
  dplyr::mutate(
    across(oc.donor_takeover:oc.rec_turnover, ~.x / n.FMT)
  ) %>%
  arrange(oc.rec_resilience) -> dat.transmission.scored.outcome.subset.summarised.by_outcome
```

Pre-compute different types of species rankings.

```{r}
#Get order of species
dat.transmission.scored.outcome.subset.summarised %>% pull(species) -> order.species.frac_rec
dat.transmission.scored.outcome.subset.summarised.by_outcome %>% pull(species) -> order.species.resilience

#Get ordered labels as well
data.phenotype.gmm.gt_20 %>%
  filter(mOTU %in% order.species.frac_rec) %>%
  dplyr::mutate(mOTU = fct_relevel(mOTU, order.species.frac_rec)) %>%
  arrange(mOTU) %>%
  pull(label) -> order.label.frac_rec

data.phenotype.gmm.gt_20 %>%
  filter(mOTU %in% order.species.resilience) %>%
  dplyr::mutate(mOTU = fct_relevel(mOTU, order.species.resilience)) %>%
  arrange(mOTU) %>%
  pull(label) -> order.label.resilience
```

### Make Plots

Generate outcome-ranked plot.

```{r, fig.width=10, fig.height=4}
dat.transmission.scored.outcome.subset.summarised %>%
  dplyr::select(-frac.don, -frac.rec, -frac.coexist) %>%
  pivot_longer(cols = c(-species, -n.FMT), names_to = "type", values_to = "fraction") %>%
  dplyr::mutate(
    fraction = case_when(
      type == "sp_lost" ~ pmax(0 - 1, 0 - fraction),
      fraction > 1 ~ 1,
      TRUE ~ fraction
    )
  ) %>%
  dplyr::mutate(
    type = fct_relevel(type, c(
      "reject_novel",
      "engraft_novel",
      "influx_novel",
      "influx_consp",
      "engraft_consp",
      "coexist_consp.donor",
      "coexist_consp.rec",
      "reject_consp",
      "sp_retained",
      "sp_lost"
    ))
  ) %>%
  left_join(data.phenotype.gmm.gt_20 %>% dplyr::select(mOTU, order), by = c("species" = "mOTU")) %>%
  ggplot(aes(x = species, y = fraction, group = type, fill = type)) +
    geom_bar(position = "stack", stat = "identity") +
    geom_hline(yintercept = c(0,1), colour = "#525252", size = 0.2) +
    scale_x_discrete(limits = order.species.frac_rec, labels = order.label.frac_rec, position = "top") +
    scale_fill_manual(values = c(
      PLOT$colour$outcome$rejection_novel,
      PLOT$colour$outcome$engraft_novel,
      PLOT$colour$outcome$influx_novel,
      PLOT$colour$outcome$influx_consp,
      PLOT$colour$outcome$engraft_consp,
      PLOT$colour$outcome$coexist.donor,
      PLOT$colour$outcome$coexist.rec,
      PLOT$colour$outcome$rejection_consp,
      PLOT$colour$outcome$species_retained,
      PLOT$colour$outcome$species_lost
    )) +
    ylab("Average Outcome post FMT") +
    theme_bw(base_family = PLOT$font$base_family, base_size = PLOT$font$size.base) +
    theme(
      axis.title = element_blank(),
      axis.text.x = element_blank(),
      #axis.text.x = element_text(angle = -45, vjust = -1, hjust = 1),
      axis.text.y = element_blank(),
      panel.grid = element_blank(),
      legend.position = "none"
    ) -> p.engraftment_frac.ranked
p.engraftment_frac.ranked
```

Do the same thing, but faceted by taxonomic group.

```{r, fig.width=10, fig.height=4}
dat.transmission.scored.outcome.subset.summarised %>%
  dplyr::select(-frac.don, -frac.rec, -frac.coexist) %>%
  pivot_longer(cols = c(-species, -n.FMT), names_to = "type", values_to = "fraction") %>%
  dplyr::mutate(
    fraction = case_when(
      type == "sp_lost" ~ pmax(0 - 1, 0 - fraction),
      fraction > 1 ~ 1,
      TRUE ~ fraction
    )
  ) %>%
  dplyr::mutate(
    type = fct_relevel(type, c(
      "reject_novel",
      "engraft_novel",
      "influx_novel",
      "influx_consp",
      "engraft_consp",
      "coexist_consp.donor",
      "coexist_consp.rec",
      "reject_consp",
      "sp_retained",
      "sp_lost"
    ))
  ) %>%
  left_join(data.phenotype.gmm.gt_20 %>% dplyr::select(mOTU, kingdom:genus, label), by = c("species" = "mOTU")) %>%
  mutate(
    taxonomy.summarised = case_when(
      genus %in% c("Bacteroides", "Phocaeicola") ~ "Bacteroides sp.",
      order == "Bacteroidales" ~ "other Bacteroidales",
      phylum %in% c("Proteobacteria") ~ as.character(phylum),
      class %in% c("Bacilli", "Negativicutes") ~ as.character(class),
      order %in% c("Actinomycetales", "Coriobacteriales") ~ as.character(order),
      grepl("Blauti", genus) ~ "Blautia sp.",
      grepl("Eubacter", genus) ~ "Eubacterium sp.",
      grepl("Dorea", genus) ~ "Dorea sp.",
      grepl("Enteroclost", genus) ~ "Enterocloster sp.",
      grepl("Lachnospira", genus) ~ "Lachnospira sp.",
      genus %in% c("Faecalibacterium") ~ "Faecalibacterium sp.",
      genus %in% c("Gemmiger") ~ "Gemmiger sp.",
      grepl("Ruminococc", genus) ~ "Ruminococcus sp.",
      grepl("Oscillibacter", genus) ~ "Oscillibacter sp.",
      genus %in% c("Lawsonibacter") ~ "Lawsonibacter sp.",
      family %in% c("Lachnospiraceae", "Ruminococcaceae", "Oscillospiraceae") ~ paste("other", as.character(family)),
      class %in% c("Clostridia", "Clostridia_A") ~ "other Clostridia",
      #order %in% c("Peptostreptococcales", "Monoglobales_A", "Christensenellales", "Oscillospirales", "Lachnospirales") ~ "other Clostridia",
      TRUE ~ "other"
    ),
    taxonomy.summarised = fct_relevel(taxonomy.summarised, c(
      "Bacteroides sp.",
      "other Bacteroidales",
      "Actinomycetales",
      "Coriobacteriales",
      "Proteobacteria",
      "Bacilli",
      "Negativicutes",
      "Oscillibacter sp.",
      "Lawsonibacter sp.",
      "other Oscillospiraceae",
      "Blautia sp.",
      "Eubacterium sp.",
      "Dorea sp.",
      "Enterocloster sp.",
      "Lachnospira sp.",
      "other Lachnospiraceae",
      "Faecalibacterium sp.",
      "Gemmiger sp.",
      "Ruminococcus sp.",
      "other Ruminococcaceae",
      "other Clostridia",
      "other"
    )),
    species = fct_relevel(species, order.species.frac_rec),
    label = fct_relevel(label, order.label.frac_rec)
  ) -> dat.outcome.stacked_bar

dat.outcome.stacked_bar %>%
  filter(type %in% c("sp_retained", "reject_consp")) %>%
  group_by(species, taxonomy.summarised) %>%
  summarise(fraction = sum(fraction)) %>%
  group_by(taxonomy.summarised) %>%
  summarise(fraction = mean(fraction)) -> tmp.summarised.persist

dat.outcome.stacked_bar %>%
  filter(type %in% c("sp_retained", "reject_consp", "coexist_consp.donor", "coexist_consp.rec")) %>%
  group_by(species, taxonomy.summarised) %>%
  summarise(fraction = sum(fraction)) %>%
  group_by(taxonomy.summarised) %>%
  summarise(fraction = mean(fraction)) -> tmp.summarised.persist_coexist

dat.outcome.stacked_bar %>%
  ggplot(aes(x = species, y = fraction, group = type, fill = type)) +
    geom_bar(position = "stack", stat = "identity") +
    geom_hline(yintercept = c(0,1), colour = "#525252", size = 0.2) +
    scale_fill_manual(values = c(
      PLOT$colour$outcome$rejection_novel,
      PLOT$colour$outcome$engraft_novel,
      PLOT$colour$outcome$influx_novel,
      PLOT$colour$outcome$influx_consp,
      PLOT$colour$outcome$engraft_consp,
      PLOT$colour$outcome$coexist.donor,
      PLOT$colour$outcome$coexist.rec,
      PLOT$colour$outcome$rejection_consp,
      PLOT$colour$outcome$species_retained,
      PLOT$colour$outcome$species_lost
    )) +
    facet_grid(. ~ taxonomy.summarised, scales = "free_x", space = "free_x") +
    geom_hline(data = tmp.summarised.persist, aes(yintercept = fraction), colour = "#cc4c02", size = 0.5, linetype = "dashed") +
    geom_hline(data = tmp.summarised.persist_coexist, aes(yintercept = fraction), colour = "#993404", size = 0.5, linetype = "dashed") +
    theme_minimal(base_family = PLOT$font$base_family, base_size = PLOT$font$size.base) +
    theme(
      axis.title = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      panel.grid = element_blank(),
      strip.text = element_blank(),
      legend.position = "none"
    ) -> p.engraftment_frac.ranked
p.engraftment_frac.ranked
```

Re-plot the fraction of recipient resilience, coexistence, donor takeover and novel strain influx events, gridded by taxonomy.

```{r, fig.height=4, fig.width=10}
dat.outcome.stacked_bar %>%
  filter(!type %in% c("sp_retained", "reject_consp", "sp_lost", "reject_novel")) %>%
  mutate(
    outcome_summary = case_when(
      type %in% c("engraft_novel", "engraft_consp", "coexist_consp.donor", "coexist_consp.rec") ~ "colonisation",
      type %in% c("influx_novel", "influx_consp") ~ "influx"
    ),
    type = fct_relevel(type, rev(c(
      "engraft_novel",
      "engraft_consp",
      "coexist_consp.donor",
      "coexist_consp.rec",
      "influx_novel",
      "influx_consp"
    )))
  ) -> tmp

tmp %>%
  filter(outcome_summary == "influx") %>%
  group_by(species, outcome_summary, taxonomy.summarised) %>%
  summarise(fraction = sum(fraction)) %>%
  group_by(outcome_summary, taxonomy.summarised) %>%
  summarise(fraction = mean(fraction)) -> tmp.summarised.influx

tmp %>%
  filter(type %in% c("engraft_novel", "engraft_consp")) %>%
  group_by(species, outcome_summary, taxonomy.summarised) %>%
  summarise(fraction = sum(fraction)) %>%
  group_by(outcome_summary, taxonomy.summarised) %>%
  summarise(fraction = mean(fraction)) -> tmp.summarised.takeover

tmp %>%
  filter(outcome_summary == "colonisation") %>%
  group_by(species, outcome_summary, taxonomy.summarised) %>%
  summarise(fraction = sum(fraction)) %>%
  group_by(outcome_summary, taxonomy.summarised) %>%
  summarise(fraction = mean(fraction)) -> tmp.summarised.colonisation

tmp %>%
  ggplot(aes(x = species, y = fraction, group = interaction(type, outcome_summary), fill = type)) +
    geom_hline(yintercept = c(0, 0.5, 1), colour = "#525252", size = 0.2) +
    geom_bar(position = "stack", stat = "identity") +
    scale_fill_manual(values = rev(c(
      PLOT$colour$outcome$engraft_novel,
      PLOT$colour$outcome$engraft_consp,
      PLOT$colour$outcome$coexist.donor,
      PLOT$colour$outcome$coexist.rec,
      PLOT$colour$outcome$influx_novel,
      PLOT$colour$outcome$influx_consp
    ))) +
    facet_grid(outcome_summary ~ taxonomy.summarised, scales = "free_x", space = "free_x") +
    geom_hline(data = tmp.summarised.colonisation, aes(yintercept = fraction), colour = "#6baed6", size = 0.5, linetype = "dashed") +
    geom_hline(data = tmp.summarised.takeover, aes(yintercept = fraction), colour = "#2171b5", size = 0.5, linetype = "dashed") +
    geom_hline(data = tmp.summarised.influx, aes(yintercept = fraction), colour = "#ae017e", size = 0.5, linetype = "dashed") +
    theme_minimal(base_family = PLOT$font$base_family, base_size = PLOT$font$size.base) +
    theme(
      axis.title = element_blank(),
      axis.text.x = element_blank(),
      #axis.text.x = element_text(angle = -45, vjust = -1, hjust = 1),
      axis.text.y = element_blank(),
      panel.grid = element_blank(),
      strip.text = element_blank(),
      legend.position = "none"
    ) -> p.engraftment_frac.ranked.subsets
p.engraftment_frac.ranked.subsets
```

Egg these plots together.

```{r, fig.width=10, fig.height=6}
require(egg)

ggarrange(
  p.engraftment_frac.ranked + theme(plot.margin = margin(r = 0, l = 0, t = 0, b = 1), panel.spacing.x = unit(0.05, "cm")),
  p.engraftment_frac.ranked.subsets + theme(plot.margin = margin(r = 0, l = 0, t = 1, b = 0), panel.spacing.x = unit(0.05, "cm")),
  heights = c(3, 2)
) -> p.engraftment_frac.ranked.combined

ggsave(p.engraftment_frac.ranked.combined, filename = paste0(PARAM$folder.results, "engraftment.ranking_species.stacked_bars.pdf"), width=15, height=10, useDingbats = F, units = "cm")

```

Plot AUCs for most common outcomes as matching lollipops.

```{r, fig.width=10, fig.height=4}
collect.res.engraft.lasso.metric %>%
  filter(metric == "auc") %>%
  filter(species %in% use.species.engraftment.gt_50) %>%
  filter(var.type == "ex_ante.combined") %>%
  filter(resp.var %in% c("oc.donor_takeover", "oc.donor_engraftment", "oc.rec_persistence", "oc.rec_turnover")) %>% group_by(resp.var) %>% summarise(mu = mean(avg), sd = sd(avg))
  select(species, resp.var, avg) %>%
  bind_rows(
    tibble(species = c("meta_mOTU_v2_6218", "meta_mOTU_v2_6288", "meta_mOTU_v2_5934", "ANI_AL_95_00021", "meta_mOTU_v2_7682", "ref_mOTU_v2_5068", "meta_mOTU_v2_6107", "ref_mOTU_v2_0144"), resp.var = "oc.donor_takeover", avg = 1)
  ) %>%
  mutate(resp.var = fct_relevel(resp.var, c("oc.rec_persistence", "oc.donor_engraftment", "oc.donor_takeover", "oc.rec_turnover"))) %>%
  mutate(avg = ifelse(avg < 0.5, 0.5, avg)) %>%
  left_join(data.phenotype.gmm.gt_20 %>% dplyr::select(mOTU, kingdom:genus, label), by = c("species" = "mOTU")) %>%
  mutate(
    taxonomy.summarised = case_when(
      genus %in% c("Bacteroides", "Phocaeicola") ~ "Bacteroides sp.",
      order == "Bacteroidales" ~ "other Bacteroidales",
      phylum %in% c("Proteobacteria") ~ as.character(phylum),
      class %in% c("Bacilli", "Negativicutes") ~ as.character(class),
      order %in% c("Actinomycetales", "Coriobacteriales") ~ as.character(order),
      grepl("Blauti", genus) ~ "Blautia sp.",
      grepl("Eubacter", genus) ~ "Eubacterium sp.",
      grepl("Dorea", genus) ~ "Dorea sp.",
      grepl("Enteroclost", genus) ~ "Enterocloster sp.",
      grepl("Lachnospira", genus) ~ "Lachnospira sp.",
      genus %in% c("Faecalibacterium") ~ "Faecalibacterium sp.",
      genus %in% c("Gemmiger") ~ "Gemmiger sp.",
      grepl("Ruminococc", genus) ~ "Ruminococcus sp.",
      grepl("Oscillibacter", genus) ~ "Oscillibacter sp.",
      genus %in% c("Lawsonibacter") ~ "Lawsonibacter sp.",
      family %in% c("Lachnospiraceae", "Ruminococcaceae", "Oscillospiraceae") ~ paste("other", as.character(family)),
      class %in% c("Clostridia", "Clostridia_A") ~ "other Clostridia",
      TRUE ~ "other"
    ),
    taxonomy.summarised = fct_relevel(taxonomy.summarised, c(
      "Bacteroides sp.",
      "other Bacteroidales",
      "Actinomycetales",
      "Coriobacteriales",
      "Proteobacteria",
      "Bacilli",
      "Negativicutes",
      "Oscillibacter sp.",
      "Lawsonibacter sp.",
      "other Oscillospiraceae",
      "Blautia sp.",
      "Eubacterium sp.",
      "Dorea sp.",
      "Enterocloster sp.",
      "Lachnospira sp.",
      "other Lachnospiraceae",
      "Faecalibacterium sp.",
      "Gemmiger sp.",
      "Ruminococcus sp.",
      "other Ruminococcaceae",
      "other Clostridia",
      "other"
    )),
    species = fct_relevel(species, order.species.frac_rec),
    label = fct_relevel(label, order.label.frac_rec)
  ) -> dat.auc.for_stacked_bars


dat.auc.for_stacked_bars %>%
  ggplot(aes(x = label, y = avg, colour = resp.var)) +
    geom_hline(yintercept = c(0.5, 0.75, 1), colour = "#525252", size = 0.2) +
    geom_point(size = 0.75) +
    geom_segment(aes(x = label, xend = label, y = avg, yend = 0.5), size = 0.2) +
    scale_colour_manual(values = c(
      PLOT$colour$outcome$species_retained,
      PLOT$colour$outcome$engraft_consp,
      PLOT$colour$outcome$engraft_novel,
      PLOT$colour$outcome$influx_novel
    )) +
    facet_grid(resp.var ~ taxonomy.summarised, space = "free_x", scales = "free_x") +
    theme_minimal() +
    theme(
      axis.title = element_blank(),
      axis.text.x = element_blank(),
      #axis.text.x = element_text(angle = -45, vjust = -1, hjust = 1),
      axis.text.y = element_blank(),
      panel.grid = element_blank(),
      strip.text = element_blank(),
      plot.margin = margin(r = 0, l = 0, t = 0, b = 0),
      panel.spacing.x = unit(0.05, "cm"),
      panel.spacing.y = unit(0.1, "cm"),
      legend.position = "none"
    ) -> p.auc.for_stacked

ggsave(p.auc.for_stacked, filename = paste0(PARAM$folder.results, "engraftment.ranking_species.auc_for_stacked_bars.pdf"), width=15, height=2.5, useDingbats = F, units = "cm")
```

Generate margin plot summarizing the frequencies for each outcome.

```{r}
size.point <- 0.2

dat.transmission.scored.outcome.subset.summarised %>%
  dplyr::select(-frac.don, -frac.rec, -frac.coexist) %>%
  pivot_longer(cols = c(-species, -n.FMT), names_to = "type", values_to = "fraction") %>%
  dplyr::mutate(
    fraction = case_when(
      fraction > 1 ~ 1,
      TRUE ~ fraction
    )
  ) %>%
  dplyr::mutate(
    type = fct_relevel(type, rev(c(
      "reject_novel",
      "engraft_novel",
      "influx_novel",
      "influx_consp",
      "engraft_consp",
      "coexist_consp.donor",
      "coexist_consp.rec",
      "reject_consp",
      "sp_retained",
      "sp_lost"
    )))
  ) %>%
  ggplot(aes(y = type, x = fraction, fill = type, colour = type)) +
    geom_vline(xintercept = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.75, 1), colour = "#969696", lwd = 0.2) +
    #Add jittered points
    geom_point(position = position_jitterdodge(jitter.width = 2), alpha = 0.7, size = size.point) +
    #Add boxplot
    geom_boxplot(outlier.colour = NA, lwd = 0.2, alpha = 1) +
    stat_summary(fun = median, geom = "crossbar", width = 0.75, colour = "#ffffff", lwd = 0.4) +
    stat_summary(fun = median, geom = "point", colour = "#ffffff", size = 1, shape = 1) +
    #Fix scales
    scale_fill_manual(values = rev(c(
      PLOT$colour$outcome$rejection_novel,
      PLOT$colour$outcome$engraft_novel,
      PLOT$colour$outcome$influx_novel,
      PLOT$colour$outcome$influx_consp,
      PLOT$colour$outcome$engraft_consp,
      PLOT$colour$outcome$coexist.donor,
      PLOT$colour$outcome$coexist.rec,
      PLOT$colour$outcome$rejection_consp,
      PLOT$colour$outcome$species_retained,
      PLOT$colour$outcome$species_lost
    ))) +
    scale_colour_manual(values = rev(c(
      PLOT$colour$outcome$rejection_novel,
      PLOT$colour$outcome$engraft_novel,
      PLOT$colour$outcome$influx_novel,
      PLOT$colour$outcome$influx_consp,
      PLOT$colour$outcome$engraft_consp,
      PLOT$colour$outcome$coexist.donor,
      PLOT$colour$outcome$coexist.rec,
      PLOT$colour$outcome$rejection_consp,
      PLOT$colour$outcome$species_retained,
      PLOT$colour$outcome$species_lost
    ))) +
    xlim(0, 1) +
    theme_minimal(base_family = PLOT$font$base_family, base_size = PLOT$font$size.base) +
    theme(
      axis.title = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid = element_blank(),
      panel.background = element_blank(),
      plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm"),
      legend.position = "none"
    ) -> p.outcome.boxplots
p.outcome.boxplots

ggsave(p.outcome.boxplots, filename = paste0(PARAM$folder.results, "engraftment.ranking_species.summary_boxplots.pdf"), width=1.45, height=3.52, useDingbats = F, units = "cm")
```


### Phenotype plots

Test marginal association of species phenotype to overall engraftment success.

```{r, fig.width=1, fig.height=1}
phenotype.variables <- c("motility", "sporulation", "oxygen.facultative", "acetate_production", "propionate_production", "butyrate_production")
phenotype_metabolism.variables <- c("saccharolysis", "proteolysis", "lipolysis")

dat.transmission.scored.engraftment %>%
  group_by(fmt.id) %>% top_n(1, timepoint.fmt) %>% ungroup() %>%
  filter(species %in% use.species.engraftment) %>%
  dplyr::mutate(
    engraftment = case_when(
      n.pos.scored < 100 ~ 0,
      !inc.donor ~ 0,
      outcome %in% c("rejection novel", "species lost") ~ 0,
      TRUE ~ donor.specific
    )
  ) %>%
  dplyr::select(species, phenotype.variables, phenotype_metabolism.variables, outcome) %>%
  dplyr::select(-acetate_production, -motility, -oxygen.facultative) %>%
  group_by(across(-outcome)) %>%
  dplyr::summarise(engraft.frac = sum(outcome %in% c("engraftment novel", "engraftment conspecific", "coexistence")) / n()) %>%
  pivot_longer(c(-species, -engraft.frac), names_to = "variable", values_to ="value") %>%
  drop_na(value) -> dat.tmp

dat.tmp %>%
  group_by(variable) %>%
  nest() %>%
  dplyr::mutate(
    model = purrr::map(data, function(df) {anova(lm(engraft.frac ~ value, data = df)) %>% broom::tidy() %>% mutate(rsq = sumsq / sum(sumsq))}),
    cor = purrr::map(data, ~cor(.$engraft.frac, .$value, method = "spearman"))
  ) %>%
  dplyr::select(-data) %>%
  unnest() %>%
  filter(term != "Residuals") -> res.engraftment.phenotype

#Plot
res.engraftment.phenotype %>%
  dplyr::mutate(rsq.directional = ifelse(cor < 0, 0 - rsq, rsq)) %>%
  dplyr::mutate(
    type = case_when(
      p.value < 0.05 & cor > 0 ~ "pos",
      p.value < 0.05 & cor < 0 ~ "neg",
      TRUE ~ "n.s."
    )
  ) %>%
  ggplot(aes(x = rsq, y = variable, colour = type)) +
    geom_vline(xintercept = 0, colour = "#252525") +
    geom_segment(aes(x = rsq, xend = 0, y = variable, yend = variable)) +
    geom_point(size = 1) +
    xlim(0, 0.1) +
    scale_color_manual(values = c("grey", "#542788", "#e08214")) +
    scale_y_discrete(limits = rev(c("propionate_production", "butyrate_production", "saccharolysis", "proteolysis", "lipolysis"))) +
    theme_bw(base_family = PLOT$font$base_family, base_size = PLOT$font$size.base) +
    theme(
      axis.title = element_blank(),
      axis.text.x = element_blank(),
      #axis.text.x = element_text(angle = -45, vjust = -1, hjust = 1),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      legend.position = "none"
    ) -> p.engraftment.phenotype.margin
p.engraftment.phenotype.margin
```



